Generalized Tractability

### 2023-04-21 GeneralizedTractability_230418.tex
Updates from Overleaf

Let $\cf_d$ and $\cg_d$ be Hilbert spaces, and $\SOL_d : \cf_d \to \cg_d$ be a linear solution operator  with adjoint $\SOL_d^*$ such that $\SOL_d^*\SOL_d : \cf_d \to \cf_d$ has eigenvalues and orthonormal eigenvectors
\begin{equation} \label{eq:eigenvalues}
\lambda_{\max}^2 \ge \lambda_{1,d}^2 \ge \lambda_{2,d}^2 \ge \cdots, \qquad u_{1,d}, u_{2,d}, \ldots, \qquad d \in \naturals.
\end{equation}
This means that the non-negative $\lambda_{i,d}$ are the singular values of the \peter{operator} $\SOL_d$. 
We make the technical assumption that $\SOL_d$ is a compact operator. This implies (see, e.g., \cite{NW08}) that the $\lambda_{i,d}$ converge to zero as $i$ tends to infinity, which in turn implies that the problem is solvable by suitable algorithms.

To avoid trivial cases, we assume that there are an infinite number of positive eigenvalues. \peter{We remark that in the literature on Information-Based Complexity the singular values of $\SOL_d$ are often denoted by $\sqrt{\lambda_{i,d}}$ instead of $\lambda_{i,d}$. However, since we assume non-negativity of the $\lambda_{i,d}$, the notation used here is equivalent.}

The goal is to find an approximate solution, $\APP_d:\cb_d \times (0,\infty) \to \cg_d$ satisfying
\begin{equation}
    \label{eq:errcrit}
    \norm[\cg_d]{\SOL_d(f) - \APP_d(f,\varepsilon)} \le \varepsilon \qquad \forall f \in \cb_d, \ \varepsilon \in (0,\infty)
\end{equation}
where $\cb_d$ is the unit ball in $\cf_d$ and $\APP_d(f,\varepsilon)$ is allowed to depend on arbitrary linear functionals \peter{(note that we also allow adaptive information, i.e., the linear functionals may sequentially depend on each other)}.

The information complexity of the problem $\SOL_d$ is denoted $\comp(\varepsilon,d)$ and defined as the minimum number of linear functionals required to satisfy the error criterion \eqref{eq:errcrit}, 
\begin{equation*}
\comp:(0,\infty) \times \naturals \to \natzero.
\end{equation*}
By definition, the complexity is non-decreasing  as $\varepsilon$ tends to zero.  We also expect it to be non-decreasing as $d$ tends to infinity.  \peter{This means that we are actually considering 
a whole sequence of problems  $\SOL_d : \cf_d \to \cg_d$, for $d\in\naturals$, and we furthermore assume} that the Hilbert spaces are defined in such a way that $\lambda_{\cdot,d}$ is non-decreasing for every $d$, as spelled out in \eqref{eq:eigenvalues}, although this assumption is not needed for our results.

@@

The question is,``How hard is this problem as $\varepsilon$ tends to zero or as $d$ tends to infinity?''  We want to bound the complexity in terms of a simple function of $\varepsilon$ and $d$ and identify simple conditions on the singular values that are equivalent to this bound holding.

>>>>

Let $\{\cf_d\}_{d \in \naturals}$ and $\{\cg_d\}_{d \in \naturals}$ be sequences of Hilbert spaces, and $\{\SOL_d : \cf_d \to \cg_d\}_{d \in \naturals}$ be a sequence of linear solution operators  with adjoints $\SOL_d^*$ such that $\SOL_d^*\SOL_d : \cf_d \to \cf_d$ has eigenvalues and orthonormal eigenvectors
\begin{equation} \label{eq:eigenvalues}
\lambda_{\max}^2 \ge \lambda_{1,d}^2 \ge \lambda_{2,d}^2 \ge \cdots, \qquad u_{1,d}, u_{2,d}, \ldots, \qquad d \in \naturals.
\end{equation}
This means that the non-negative $\lambda_{i,d}$ are the singular values of the operator $\SOL_d$. We remark that in the literature on Information-Based Complexity the singular values of $\SOL_d$ are often denoted by $\sqrt{\lambda_{i,d}}$ instead of $\lambda_{i,d}$. However, since we assume non-negativity of the $\lambda_{i,d}$, the notation used here is equivalent.  The non-decreasing ordering of the eigenvalues for each $d$ facilitates the proof of conditions for tractability, but our conditions do not depend on the ordering.

We make the technical assumption that $\SOL_d$ is a compact operator. This implies (see, e.g., \cite{NW08}) that $\lambda_{i,d}$ converges to zero as $i$ tends to infinity for every $d$, which in turn implies that the problem is solvable by suitable algorithms. To avoid trivial cases, we assume that there are an infinite number of positive eigenvalues.

The goal is to find a sequence of approximate solution operators, $\{\APP_d:\cb_d \times (0,\infty) \to \cg_d\}_{d \in \naturals}$ satisfying
\begin{equation}
    \label{eq:errcrit}
    \norm[\cg_d]{\SOL_d(f) - \APP_d(f,\varepsilon)} \le \varepsilon \qquad \forall f \in \cb_d, \ \varepsilon \in (0,\infty)
\end{equation}
where $\cb_d$ is the unit ball in $\cf_d$ and $\APP_d(f,\varepsilon)$ is allowed to depend on arbitrary linear functionals.  Note that we also allow adaptive information, i.e., the linear functionals may sequentially depend on each other.

The information complexity of the problem $\SOL_d$ is given by the function $\comp:(0,\infty) \times \naturals \to \natzero$, where $\comp(\varepsilon,d)$ is the minimum number of linear functionals required to satisfy the error criterion \eqref{eq:errcrit}.
By definition, the complexity is non-decreasing  as $\varepsilon$ tends to zero.  We also expect it to be non-decreasing as $d$ tends to infinity, since $d$ is typically the number of variables for the functions in $\cf_d$.

@@

The question of tractability is one of determining how fast the complexity increases as $\varepsilon$ tends to zero and/or as $d$ tends to infinity.  We want to bound the complexity in terms of a simple function of $\varepsilon^{-1}$ and $d$ and identify simple conditions on the singular values that are equivalent to this bound holding.
